---
params:
 title: ""
 subtitle: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::pdf_document2: default
  bookdown::html_document2:
    toc: true
    toc_float: true
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace # <- don't do this - rmeoves params set in yaml!

# Set start time ----
startTime <- proc.time()

library(nzGREENGridDataR)

# run gg set up
nzGREENGridDataR::setup() # sets a load of parameters which are then accessible via ggrParams$xxx

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # writing to files
             "Hmisc", # for describe
             "knitr" # for kable
)
# load them
nzGREENGridDataR::loadLibraries(rmdLibs)


# Local paramaters
ggrParams$outPath <- "/Volumes/hum-csafe/Research Projects/GREEN Grid/Clean_data/safe/survey/"
```

\newpage

# About

## Report circulation:

 * Public - this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r lubridate::year(today())`) `r params$title` `r params$subtitle`, `r ggrParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

```{r includeHistory, child=ggrParams$historyGenericRmd}
```
 
Specific history of this code:

 * https://github.com/dataknut/nzGREENGridDataR/commits/master/dataProcessing/surveys/createHouseholdAttributes.Rmd

## Requirements:

This report uses the original nz GREEn grid household survey and meta data.

## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

The prupose of this report is to: 

 * create a household attribute file that can be linked to the project power monitoring data;
 * describe the household attribute data.

The resulting cleaned data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners.

The data contains a unique household id which can be used to link it to the NZ GREEN Grid time use diaries and dwelling/appliance surveys. With some additional non-disclosure checks it should also be safe to archive all of these linkable datasets for re-use via the UK [reshare](http://reshare.ukdataservice.ac.uk/) service.

# Load data

In this section we load metadata from `r ggrParams$gsHHMasterFile`.

```{r load metadata}
hhAttributesDT <- nzGREENGridDataR::getHouseholdData(ggrParams$gsHHMasterFile)
t <- with(hhAttributesDT, table(Location, useNA = "always"))
kable(caption = "Sample location", t)
```

In total we have `r nrow(hhAttributesDT)` households in two sample areas.

## Data description

```{r summarise metadata}
Hmisc::describe(hhAttributesDT)
```


# Describe data

NA usually means not known.

## Number of adults

```{r nAdults}
t <- with(hhAttributesDT, table(nAdults, Location, useNA = "always"))
knitr::kable(caption = "Number of adults in household by location", t)
```

## Number of teenagers

```{r nTeens}
t <- with(hhAttributesDT, table(nTeenagers13_18, Location, useNA = "always"))
knitr::kable(caption = "Number of teenagers in household by location", t)
```

## Number of children

```{r nChildren}
t <- with(hhAttributesDT, table(nChildren0_12, Location, useNA = "always"))
knitr::kable(caption = "Number of children in household by location", t)
```

## Notes

These may have been set for any number of reasons and mean the monitoring data should be used with caution.

```{r outliers}
t <- with(hhAttributesDT, table(notes, Location, useNA = "always"))
knitr::kable(caption = "Notes by location", t)
```

# Summary

```{r saveData}
ofile <- paste0(ggrParams$outPath, "ggHouseholdAttributes.csv")
readr::write_csv(hhAttributesDT, ofile)
```

The cleaned data for the `r nrow(hhAttributesDT)` households has been saved as a .csv file to:

 * `r ggrParams$outPath` 

The following table shows the key columns of the household attributes file. The data can be linked to the gridSpy data using hhID.

The purpose of the newID is to enable the flagging of re-used grid spy units. As an example unit rf_15 was re-used in a different household. Data users should therefore make sure that the correct household data (rf_15a or rf_15b) is linked to the grid spy data (coded rf_15) at the correct date.

Note also that data exists in the grid spy power demand data for whom no household data exists (e.g. rf_01 & rf_02).

```{r data header}
kable(caption = "Household data: key columns", hhAttributesDT[, c("hhID", "Location", "nAdults", "newID", "r_stopDate", "notes")][order(hhID)])
```


# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for csv reading/writing [@readr]
 * Hmisc - for describe [@Hmisc]
 * knitr - to create this document & neat tables [@knitr]
 * readxl - reading .xlsx [@readxl]
 * nzGREENGridDataR - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
