---
params:
 title: ""
 subtitle: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace # <- don't do this - rmeoves params set in yaml!

# Set start time ----
startTime <- proc.time()

library(nzGREENGridDataR)

# run gg set up
nzGREENGridDataR::setup() # sets a load of parameters which are then accessible via ggrParams$xxx

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # reading/writing csv files
             "readxl", # reading xlsx
             "Hmisc", # for describe
             "kableExtra" # for kable
)
# load them
nzGREENGridDataR::loadLibraries(rmdLibs)


# Local paramaters
ggrParams$outPath <- "/Volumes/hum-csafe/Research Projects/GREEN Grid/Clean_data/safe/survey/"
  
```

\newpage

# About

## Report circulation:

 * Public - this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r lubridate::year(today())`) `r params$title` `r params$subtitle`, `r ggrParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

```{r includeHistory, child=ggrParams$historyGenericRmd}
```
 
Specific history of this report:

 * https://github.com/dataknut/nzGREENGridDataR/commits/master/dataProcessing/surveys/createHouseholdAttributes.Rmd

## Requirements:

This report uses the original nz GREEN Grid household survey and meta data.

## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
The prupose of this report is to: 

 * create a household attribute file that can be linked to the project power monitoring data;
 * describe the household attribute data.

The resulting cleaned data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners.

The data contains a unique household id (`linkID`) which can be used to link it to the NZ GREEN Grid time use diaries and dwelling/appliance surveys. With some additional non-disclosure checks it should also be safe to archive all of these linkable datasets for re-use via the UK [reshare](http://reshare.ukdataservice.ac.uk/) service.

# Load data

## Household master dataset
Table \@ref(tab:loadMasterData) shows the number of households in each area. 

```{r loadMasterData}
hhMasterDT <- nzGREENGridDataR::getHouseholdData(ggrParams$gsHHMasterFile) # cleans it as it loads
setkey(hhMasterDT, linkID)
t <- with(hhMasterDT, table(Location, useNA = "always"))
kableExtra::kable(caption = "Sample location (master file)", t)
```

In total we have `r nrow(hhMasterDT)` households in two sample areas.

## Appliance data

Household appliance ownership was recorded during recruitment using a detailed survey. However this data is not readily available for all households at present. Table \@ref(tab:applianceTable) shows the number of households in each area for whom summary appliance data exists. 

```{r loadApplianceData}
hhAppliancesDT <- data.table::as.data.table(readxl::read_xlsx(ggrParams$ApplianceData))
```

```{r applianceTable}
hhAppliancesDT$sample <- NULL # too detailed
setkey(hhAppliancesDT, linkID)
t <- with(hhAppliancesDT[hhMasterDT], table(Location, hasApplianceSummary, useNA = "always"))
kableExtra::kable(caption = "Sample (appliance summary file, NA indicates no data)", t)
```

The appliances recorded in this summary are shown in Table \@ref(tab:applianceNames).

```{r applianceNames}
xcols <- c("linkID", "hasApplianceSummary")
al <- names(hhAppliancesDT[, -..xcols])
kableExtra::kable(caption = "Appliances recorded", al)
```

## Household survey data

The Energy Cultures 2 survey [@ec2Survey2015] was used to collect data on household energy and transport attitudes/behaviours and usages for one of the regional samples only. The long form of the survey was used for some households and the short form for others.

```{r loadEc2Surveys, include=FALSE}
hhEc2ShortDT <- data.table::as.data.table(readr::read_csv(ggrParams$UnisonShortSurveyData))
hhEc2LongDT <- data.table::as.data.table(readr::read_csv(ggrParams$UnisonLongSurveyData))
```

Table \@ref(tab:loadEc2Survey) shows the response rates to these surveys.

```{r loadEc2Survey}
# short ----

keepShortCols <- c("linkID","hasShortSurvey",
              "StartDate",
              "Q7",
              "Q53_1",
              "Q53_2",
              "Q53_3",
              "Q53_4",
              "Q53_5",
              "Q53_6",
              "Q53_7",
              "Q54_1",
              "Q54_2",
              "Q54_3",
              "Q54_4",
              "Q54_5",
              "Q54_6",
              "Q54_7",
              "Q57",
              "Q4",
              "Q55")
keephhEc2ShortDT <- hhEc2ShortDT[!is.na(linkID), ..keepShortCols]

# long ----
keepLongCols <- c("linkID", "hasLongSurvey",
              "StartDate",
              "Q5",
              "Q7",
              "Q10#1_1_1_TEXT",
              "Q10#1_1_2_TEXT",
              "Q10#1_2_1_TEXT",
              "Q10#1_2_2_TEXT",
              "Q10#1_3_1_TEXT",
              "Q10#1_3_2_TEXT",
              "Q10#1_4_1_TEXT",
              "Q10#1_4_2_TEXT",
              "Q10#1_5_1_TEXT",
              "Q10#1_5_2_TEXT",
              "Q10#1_6_1_TEXT",
              "Q10#1_6_2_TEXT",
              "Q10#1_7_1_TEXT",
              "Q10#1_7_2_TEXT",
              "Q10#1_8_1_TEXT",
              "Q10#1_8_2_TEXT",
              "Q11_1",
              "Q14_1",
              "Q15_1",
              "Q17_1",
              "Q18_1",
              "Q30_1",
              "Q33_1",
              "Q53_1",
              "Q53_2",
              "Q53_3",
              "Q53_4",
              "Q53_5",
              "Q53_6",
              "Q53_7",
              "Q54_1",
              "Q54_2",
              "Q54_3",
              "Q54_4",
              "Q54_5",
              "Q54_6",
              "Q54_7",
              "Q57",
              "Q4",
              "Q55")
keephhEc2LongDT <- hhEc2LongDT[!is.na(linkID), ..keepLongCols] # no point keeping unknown IDs

setkey(keephhEc2ShortDT, linkID)
setkey(keephhEc2LongDT, linkID)
hhEc2CombinedDT <- rbind(keephhEc2LongDT,keephhEc2ShortDT, fill=TRUE)

# check for duplicates
# nrow(hhEc2CombinedDT)
# uniqueN(hhEc2CombinedDT$linkID)

# set date ----
hhEc2CombinedDT <- hhEc2CombinedDT[, r_startDate := lubridate::dmy_hms(StartDate)]
hhEc2CombinedDT$StartDate <- NULL
setkey(hhEc2CombinedDT, linkID)

# create combined DT
hhAttributesDT <- hhEc2CombinedDT[hhAppliancesDT][hhMasterDT]

t <- hhAttributesDT[, .(nHouseholds = .N), keyby = .(Location, hasLongSurvey, hasShortSurvey)]
kableExtra::kable(caption = "Survey responses (NA indicates no survey)", t)
```

# Final household attribute data description

Table \@ref(tab:loadEc2Survey) reports the final household attribute variables and their internal distributions. The question labels are found in Table \@ref(tab:loadEc2Labels) below.

```{r summariseHhAttributes}
Hmisc::describe(hhAttributesDT)
```


# Household attribute tables

This section reports tables of the key household attributes by sample location. Note that NA usually means not known.

## Number of adults

```{r nAdults}
t <- with(hhAttributesDT, table(nAdults, Location, useNA = "always"))
kableExtra::kable(caption = "Number of adults in household by location", t)
```

## Number of teenagers

```{r nTeens}
t <- with(hhAttributesDT, table(nTeenagers13_18, Location, useNA = "always"))
kableExtra::kable(caption = "Number of teenagers in household by location", t)
```

## Number of children

```{r nChildren}
t <- with(hhAttributesDT, table(nChildren0_12, Location, useNA = "always"))
kableExtra::kable(caption = "Number of children in household by location", t)
```

## Notes

These may have been set for any number of reasons and mean the monitoring data should be used with caution.

```{r outliers}
t <- with(hhAttributesDT, table(notes, Location, useNA = "always"))
kableExtra::kable(caption = "Notes by location", t)
```

# Summary

```{r saveData}
ofile <- paste0(ggrParams$outPath, "ggHouseholdAttributes.csv")
readr::write_csv(hhAttributesDT, ofile)
```

The cleaned data for the `r nrow(hhAttributesDT)` households has been saved as a .csv file to:

 * `r ggrParams$outPath` 

The following table shows the key columns of the household attributes file. The data can be linked to the gridSpy data using `linkID`.

The purpose of the linkID is to enable the flagging of [re-used grid spy units](gridSpy1mProcessingReport.html#41_recoding_re-allocated_grid_spy_units). As an example unit rf_15 was re-used in a different household. We have created `linkID` so that the correct household data (rf_15a or rf_15b) can be linked to the grid spy data (coded rf_15) at the correct date.

Note also that data exists in the grid spy power demand data for whom no household data exists (e.g. rf_01 & rf_02).

```{r data header}
kable(caption = "Household data: key columns", hhAttributesDT[, c("hhID", "linkID", "Location", "nAdults", "r_stopDate", "notes")][order(hhID)])
```


# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [kableExtra](https://cran.r-project.org/package=kableExtra) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for csv reading/writing [@readr]
 * Hmisc - for describe [@Hmisc]
 * kableExtra - to create neat tables [@kableExtra]
 * readxl - reading .xlsx [@readxl]
 * nzGREENGridDataR - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# Energy Cultures 2 Long Survey Questions
Table \@ref(tab:loadEc2Labels) lists the full Energy Cultures 2 survey [@ec2Survey2015] as used in the Unison sample. Only a subset of these variables are currently available in the GREEN Grid safe data package. Requests to add further variables should be made via a [github repo issue](https://github.com/dataknut/nzGREENGridDataR/labels/variableRequest). 

```{r loadEc2Labels}
ec2LabsDF <- readxl::read_xlsx(ggrParams$ec2LongSurveyLabels, sheet = "ec2Long")

kable(caption = "Energy Cultures 2 Survey labels (long)", ec2LabsDF)
```

# References
