---
params:
 title: "GREEN Grid Data Processing"
 subtitle: "Create Household Attributes File"
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::pdf_document2: default
  bookdown::html_document2:
    toc: true
    toc_float: true
bibliography: '`r paste0(nzGREENGrid::findParentDirectory("nzGREENGrid"), "/bibliography.bib")`'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace # <- don't do this - rmeoves params set in yaml!

# Set start time ----
startTime <- proc.time()

library(nzGREENGrid)

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # writing to files
             "Hmisc", # for describe
             "knitr" # for kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# run gg set up
nzGREENGrid::setup() # sets a load of parameters which are then accessible via ggParams$xxx

# Local paramaters
ggParams$outPath <- "/Volumes/hum-csafe/Research Projects/GREEN Grid/Clean_data/safe/survey/"
```

\newpage

# Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) `r params$title`: `r params$subtitle`, `r ggParams$pubLoc`.

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# About

## Circulation

```{r generic circulation, child=ggParams$circulationGenericRmd}
```

## History

```{r includeHistory, child=ggParams$historyGenericRmd}
```
 
Specific history of this code:

 * https://git.soton.ac.uk/ba1e12/nzGREENGrid/tree/master/analysis/ev

## Support

```{r includeSupport, child=ggParams$supportGenericRmd}
```

\newpage

# Introduction

The prupose of this report is to: 

 * create a household attribute file that can be linked to the project power monitoring data

The resulting cleaned data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners.

The data contains a unique household id which can be used to link it to the NZ GREEN Grid time use diaries and dwelling/appliance surveys. With some additional non-disclosure checks it should also be safe to archive all of these linkable datasets for re-use via the UK [reshare](http://reshare.ukdataservice.ac.uk/) service.

# Requirements:

 * GREEN Grid household suveys and metadata files

# Load data

In this section we load metadata from `r ggParams$gsMasterFile`.

```{r load metadata}
metaDT <- nzGREENGrid::getMetaData(ggParams$gsMasterFile)
with(metaDT, table(Location, useNA = "always"))
```

In total we have `r nrow(metaDT)` households in two sample areas.

## Data description

```{r summarise metadata}
Hmisc::describe(metaDT)
```


# Describe data

NA usually means not known.

## Number of adults

```{r nAdults}
t <- with(metaDT, table(nAdults, Location, useNA = "always"))
knitr::kable(caption = "Number of adults in household by location", t)
```

## Number of teenagers

```{r nTeens}
t <- with(metaDT, table(nTeenagers13_18, Location, useNA = "always"))
knitr::kable(caption = "Number of teenagers in household by location", t)
```

## Number of children

```{r nChildren}
t <- with(metaDT, table(nChildren0_12, Location, useNA = "always"))
knitr::kable(caption = "Number of children in household by location", t)
```

## Outlier flag

These may have been set for any number of reasons and mean the monitoring data should be used with caution.

```{r outliers}
t <- with(metaDT, table(outlierFlag, Location, useNA = "always"))
knitr::kable(caption = "Outlier flags by location", t)
```

# Summary

```{r saveData}
ofile <- paste0(ggParams$outPath, "ggHouseholdAttributes.csv")
readr::write_csv(metaDT, ofile)
```

The cleaned data for the `r nrow(metaDT)` households has been saved as a .csv file to:

 * `r ggParams$outPath` 

The following table shows the first few rows of the household attributes file:

```{r data header}
kable(caption = "Example data rows", head(metaDT))
```

The data can be linked to the gridSpy data using hhID and/or newID.

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for csv reading/writing [@readr]
 * Hmisc - for describe [@Hmisc]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
