---
params:
  hhID: "rf_38" # <- change this to switch household. That's it :-)
  title: 'NZ GREEN Grid project example:'
  subtitle: 'Testing power demand: '
title: '`r params$title`'
subtitle: '`r paste0(params$subtitle, params$hhID)`'
author: "Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    self_contained: no
    toc: yes
    toc_depth: 2
    toc_float: TRUE
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()


# Load nzGREENGrid package ----
library(nzGREENGridDataR) # local utilities

nzGREENGridDataR::setup()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "lubridate", # date & time stuff
             "ggplot2", # for fancy graphs
             "readr", # for reading & parsing .csv files
             "knitr" # for kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# Local parameters ----
# change this to suit your data location & to use a different household
gsFile <- paste0(ggrParams$dataLoc, "Clean_data/safe/gridSpy/1min/data/", params$hhID,"_all_1min_data.csv.gz")
hhFile <- paste0(ggrParams$dataLoc, "Clean_data/safe/survey/ggHouseholdAttributes.csv")

# Local functions ----


```

\newpage

# About

## Report circulation:

 * Public - this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r lubridate::year(today())`) `r params$title` `r paste0(params$subtitle, params$hhID)` `r ggrParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

Code history is generally tracked via our github [repo](https://github.com/dataknut/nzGREENGridDataR):

 * [Report history](https://github.com/dataknut/nzGREENGridDataR/commits/master/examples/testHouseholdPower.Rmd)
 * [General issues](https://github.com/dataknut/nzGREENGridDataR/issues)
 
## Requirements:

This report uses the safe version of the grid spy 1 minute data which has been processed using https://github.com/dataknut/nzGREENGridDataR/blob/master/dataProcessing/gridSpy/processGridSpy1mData.R.

## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
This report provides summary analysis of one household as an example.

# Load `r params$hhID` data

The data used to generate this report is:

 * `r gsFile`
 * `r hhFile`

```{r loadAllData}
hhDT <- data.table::as.data.table(readr::read_csv(hhFile)) # load hh data

t <- hhDT[hhID == params$hhID]

knitr::kable(caption = paste0("Summary of household data for", params$hhID), t)

gsDT <- data.table::as.data.table(readr::read_csv(gsFile, 
                                                  col_types = cols(hhID = col_character(),
                                                                   dateTime_orig = col_character(), # <- this is crucial otherwise readr attempts to parse this as a dateTime and FAILS (see https://github.com/dataknut/nzGREENGridDataR/issues/2)
                                                                   TZ_orig = col_character(),
                                                                   r_dateTime = col_datetime(format = ""),
                                                                   circuit = col_character(),
                                                                   powerW = col_double()
                                                                   )
                                                  )
                                  ) # load gs data

t <- summary(gsDT)

knitr::kable(caption = paste0("Summary of grid spy data for", params$hhID), t)

gsDT <- gsDT[, month := lubridate::month(r_dateTime, label = TRUE)]
gsDT <- gsDT[, year := lubridate::year(r_dateTime)]
gsDT <- gsDT[, obsHour := lubridate::hour(r_dateTime)]
gsDT <- gsDT[, obsDate := lubridate::date(r_dateTime)]
gsDT <- gsDT[, obsTime := hms::as.hms(r_dateTime)]
```

The first table shows household attributes such as how many people live in this household.

The second table shows a summary of the grid spy 1 minute power data. Note that:

 * the original dataTime (`dateTime_orig`) and TZ (`TZ_orig`) have been retained so that the user can check for parsing errors (see https://github.com/dataknut/nzGREENGridDataR/issues/2) if required;
 * r_datetime is the correct dateTime of each observation in UTC and will have loaded as your local timezone. If you are conducting this analysis outside NZ then you will get strange results until you use [lubridate](https://lubridate.tidyverse.org/) to tell R to use tz = "Pacific/Auckland" with this variable;

# Plot monthly power profiles

This section plots overall mean power (W) per minute per month for each circuit to show:

 * patterns of missing data (no lines)
 * patterns of consumption
 * possible dateTime issues (where consumption patterns seem to be stangely shifted in time)
 * possible PV installation

```{r plotProfiles, fig.height=8}
plotDT <- gsDT[, .(meanW = mean(powerW)), keyby = .(circuit, month, year, obsTime)
             ] # aggregate by circuit to preserve unique circuit labels in households 
# (e.g. rf_46) where names are re-used but with different ids. see ?fixCircuitLabels_rf_46
vLineAlpha <- 0.4
vLineCol <- "#0072B2" # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
timeBreaks <- c(hms::as.hms("04:00:00"), 
                hms::as.hms("08:00:00"),
                hms::as.hms("12:00:00"),
                hms::as.hms("16:00:00"),
                hms::as.hms("20:00:00"),
                hms::as.hms("24:00:00")
)

myPlot <- ggplot2::ggplot(plotDT, aes(x = obsTime, y = meanW/1000, colour = circuit)) +
  geom_line() + 
  facet_grid(month  ~ year) + 
  theme(strip.text.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  theme(legend.position = "bottom") + 
  labs(title = paste0("Montly mean power profiles by circuit plot: ", params$hhID),
       y = "Mean kW", 
       caption = paste0("gridSpy data from ", min(gsDT$r_dateTime), 
                        " to ", max(gsDT$r_dateTime),
                        "\nobsTime = Pacific/Auckland"))

myPlot + 
  scale_x_time(breaks = timeBreaks) +
  geom_vline(xintercept = timeBreaks, alpha = vLineAlpha, colour = vLineCol)
```


# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * knitr - to create this document & neat tables [@knitr]
 * nzGREENGridDataR - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
