---
params:
  circuitPattern: ""
  dateFrom: ""
  dateTo: ""
  iFile: ""
title: '`r paste0("NZ GREEN Grid Household Power Demand Profiles: ", params$circuitPattern, " (", params$dateFrom, " to ", params$dateTo, ")")`'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  pdf_document2:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document2:
    code_folding: hide
    fig_caption: true
    keep_md: true
    number_sections: true
    self_contained: no
    toc: true
    toc_float: true
    toc_depth: 2
bibliography: '`r paste0(findParentDirectory("GREENGrid"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```

```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()

library(GREENGridData) # local utilities

# Set parameters etc ----
GREENGrid::setup()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "dplyr", # data munching
             "ggplot2", # for fancy graphs
             "readr", # writing to files
             "lubridate", # for today
             "kableExtra" # for extra kable
)
# load them
loadLibraries(rmdLibs)

# Local functions ---
labelProfilePlot <- function(plot){
  # adds neat labels to X axis
  plot <- plot + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    scale_x_time(breaks = c(hms::as.hms("00:00:00"),
                            hms::as.hms("04:00:00"), hms::as.hms("08:00:00"), 
                            hms::as.hms("12:00:00"), hms::as.hms("16:00:00"),
                            hms::as.hms("20:00:00"), hms::as.hms("24:00:00")))

  return(plot)
}

# Local parameters ----
plotCaption <- paste0("Data source: ", fpath,
                      "\n", params$circuitPattern, " (", params$dateFrom, " to ", params$dateTo, ")" )
```

\newpage

# About

## Report circulation:

 * Restricted to: [NZ GREEN Grid](https://www.otago.ac.nz/centre-sustainability/research/energy/otago050285.html) project partners and contractors.

## License

```{r ccby license, child=ggParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) `r paste0("NZ GREEN Grid Household Power Demand Profiles: ", params$circuitPattern, " (", params$dateFrom, " to ", params$dateTo, ")")`, University of Otago: Dunedin, NZ.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

Code history is generally tracked via our git.soton [repo](https://github.com/CfSOtago/GREENGrid):

 * [Report history](https://github.com/CfSOtago/GREENGrid/commits/master/analysis/ggDemandProfiles)
 
## Requirements:

 * Cleaned and safe grid spy 1 minute data processed using the [nzGREENGridDataR package](https://github.com/CfSOtago/GREENGridData/tree/master/dataProcessing/gridSpy)
 * `r params$circuitPattern` circuit data pre or newly extracted from the cleaned safe data using https://github.com/CfSOtago/GREENGridData/commits/master/examples/extractCleanGridSpy1minCircuits.R

## Support

```{r generic support, child=ggParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggParams$sampleGeneric}
```

This report is intended to: 

 * load a data file containing pre-extracted `r params$circuitPattern` circuits (via string match on their circuit label) for the period `r params$dateFrom` to `r params$dateTo` (inclusive);
 * build and plot exploratory demand profiles for `r params$circuitPattern` circuits;
 * save aggregated demand profiles for `r params$circuitPattern` circuits to `r paste0(outPath, "profiles/")`.

# Load data files

## Household attribute data

In this section we load metadata from `r ggParams$hhAttributes` to link to the power data.

```{r load hhAttributes}

hhAttributesDT <- data.table::fread(ggParams$hhAttributes)


setkey(hhAttributesDT, hhID)

kableExtra::kable(caption = "Example household attribute data for sample", head(hhAttributesDT))
```

## Grid Spy data

```{r setFile}
iFile <- paste0(params$iFile, ".gz") # don't forget .gz!
#iFile <- paste0(iFile, ".gz")
```

In this section we load the pre-extracted `r params$circuitPattern` data from `r dateFrom` to `r dateTo` from `r iFile`.

```{r load gridSpy data}
# Loads pre-extracted circuit data using readr as the data is .csv.gz
gs1MinDT <- data.table::as.data.table(
  readr::read_csv(iFile,
                  col_types = cols(
                        hhID = col_character(),
                        r_dateTime = col_datetime(format = ""),
                        circuit = col_character(),
                        powerW = col_double() # <- crucial otherwise readr infers integers for some reason
                      )
                  )
)
```

The following table shows the range of different circuit labels that have been included.

```{r test circuit labels}
t <- table(gs1MinDT$circuit)

kableExtra::kable(caption = paste0("Counts of ", circuitPattern, " observations by label"), t)
```

Note that some households may have more than one `r circuitPattern` circuit as the following table indicates.

```{r test household circuits}
t <- gs1MinDT[, .(nCircuits = uniqueN(circuit)), keyby = .(hhID)]

kableExtra::kable(caption = paste0("Counts of ", circuitPattern, " labels by household"), t)
```

### Re-coding re-used monitors

The following table shows the grid spy monitors which were re-used together with their dates.

```{r testReuse}
t <- hhAttributesDT[!is.na(notes), .(notes = unique(notes)), keyby = .(hhID, linkID, r_stopDate)]

kableExtra::kable(caption = paste0("Counts of ", circuitPattern, " labels by household"), t)
```

In order to correctly link household attribute and grid spy data we therefore need to set linkID in the `r circuitPattern` data for the appropriate dates.

```{r addlinkID}
# set merge keys
setkey(hhAttributesDT, linkID)
setkey(gs1MinDT, linkID)

# foreach hhID in hhAttributes
# if !is.na(stopped)

```


## Test `r circuitPattern` data

Figure \@ref(fig:nObsTile) shows loaded data observation plots - just to confirm what `r circuitPattern` data we have. 

```{r nObsTile, fig.cap="Number of observations per household per day"}

plotDT <- gs1MinDT[, .(nObs = uniqueN(r_dateTime)), keyby = .(linkID, date = as.Date(r_dateTime))]
setkey(plotDT, linkID)

dt <- merge(plotDT, hhAttributesDT, allow.cartesian=TRUE) # not sure why we need to do this - shouldn't be duplicate hhIDs?

myPlot <- ggplot2::ggplot(dt, aes( x = as.Date(date), y = linkID, 
                               fill = nObs)) +
  geom_tile() +
  scale_fill_gradient(low = "red", high = "green") +
  scale_x_date(date_labels = "%Y %b", date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5,
                                   hjust = 0.5)) + 
  labs(caption = plotCaption,
       x = "Date"
  )

myPlot + 
  facet_grid(sample ~ .) +
  annotate("rect")

```

Figure \@ref(fig:nObsPoint) shows the same data but as a dot plot to highlight those households and dates where we did not receive 60 * 24 = 1440 observations per day.

```{r nObsPoint, fig.cap="Number of observations per household per day"}
# point plot ----
myPlot <- ggplot2::ggplot(dt[!is.na(nObs)], aes( x = date, 
                               y = nObs, 
                               colour = linkID)) +
  geom_point() +
  #facet_wrap(sample ~ .) +
  scale_x_date(date_labels = "%Y %b", date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, 
                                   hjust = 0.5)) + 
  labs(caption = plotCaption,
       x = "Date"
  )
myPlot + facet_grid(sample ~ .)
```

The following table shows the min/max observations per day and min/max dates for each household. As above, we should not see:

 * dates before 2014 or in to the future (indicates date conversion errors)
 * more than 1440 observations per day (indicates potentially duplicate observations)
 * non-integer counts of circuits as it suggests some column errors
 
 We should also not see NA in any row (indicates date conversion errors). 
 
 If we do see any of these then we still have data cleaning work to do!

```{r summaryTable}
# Stats table (so we can pick out the dateTime errors)
setkey(gs1MinDT, hhID)
mergedDT <- merge(gs1MinDT, hhAttributesDT,  allow.cartesian=TRUE)

dt <- mergedDT[, .(nObs = .N,
             minDate = min(r_dateTime, na.rm = TRUE),
             maxDate = max(r_dateTime, na.rm = TRUE)),
         keyby = .(hhID, sample)]

kableExtra::kable(caption = paste0("Summary observation stats by hhID (sorted by date last heard from) for: ", circuitPattern), dt)
```

Finally we show the total number of households which we think we have `r circuitPattern` data for.

```{r liveDataHouseholds}
plotDT <- mergedDT[, .(nHH = uniqueN(hhID)), keyby = .(date = as.Date(r_dateTime), sample)]

plotDT <- plotDT[sample == "Unison", sample := "Sample 1: Unison" ]
plotDT <- plotDT[sample == "Powerco", sample := "Sample 2: PowerCo" ]

# point plot ----
myPlot <- ggplot2::ggplot(plotDT, aes( x = date, y = nHH, fill = sample)) +
  geom_col() +
  scale_x_date(date_labels = "%Y %b", date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 0.5)) + 
  labs(caption = plotCaption,
       x = "Date"
  )
myPlot

```

The following table summarises the `r circuitPattern` data. Any surprises?

```{r summary of cols, echo = TRUE}
t <- summary(gs1MinDT)
kableExtra::kable(caption = paste0("Summary of ", circuitPattern, " circuits"), t)

```

We seem to have some negative powerW values and at least one very large power value.

Nasty surprises often lurk in histograms... The following histogram shows all observations.

```{r histo full}
ggplot2::ggplot(gs1MinDT, aes(x = powerW)) +
  geom_histogram(binwidth = 10) +
  labs(caption = plotCaption)
```

The next shows the histogram for powerW < 1000W...

```{r histo power under 1000}
ggplot2::ggplot(gs1MinDT[powerW < 1000], aes(x = powerW)) +
  geom_histogram(binwidth = 10) +
  labs(caption = plotCaption)
```

> There are a lot of zeros (as we'd expect) but why are there negative values?

# `r circuitPattern` profiles

This section produces the profiles as one for each HH but averaged over each season. Data is kept at 1 minute intervals. Note definition of season below...

```{r add season, echo = TRUE}
# add season
gs1MinDT <- gs1MinDT[, month := lubridate::month(r_dateTime, label = TRUE)]
gs1MinDT <- gs1MinDT[, season := "Summer"]
gs1MinDT <- gs1MinDT[, season := ifelse(month == "Mar" |
                                              month == "Apr" |
                                              month == "May", "Autumn", season)]
gs1MinDT <- gs1MinDT[, season := ifelse(month == "Jun" |
                                              month == "Jul" |
                                              month == "Aug", "Winter", season)]
gs1MinDT <- gs1MinDT[, season := ifelse(month == "Sep" |
                                              month == "Oct" |
                                              month == "Nov", "Spring", season)]
```

## Profile plots: means per household

Figure \@ref{fig:meanHhPlot} shows the mean power for circuits containing "`r circuitPattern`" by season.

```{r meanHhPlot, fig.cap="Mean power demand (W) per household", fig.height=6}
# add onsHourMin
gs1MinDT <- gs1MinDT[, obsHourMin := hms::as.hms(r_dateTime)]

# create plot table
plotDT <- gs1MinDT[!is.na(season), 
                     .(meanW = mean(powerW)), keyby = .(hhID, obsHourMin, season)]
setkey(plotDT, hhID)
nHHs <- uniqueN(plotDT$hhID)
myPlot <- ggplot2::ggplot(plotDT, aes(y = meanW, x = obsHourMin, colour = hhID)) +
  geom_point() +
  facet_grid(season ~ .) +
  labs(caption = paste0(plotCaption,"\nn households = ", nHHs),
       x = "Time of Day")
labelProfilePlot(myPlot)

# save the plot at high def
plotFile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_byHouseholdSeasonalProfilePlot.pdf")
ggplot2::ggsave(plotFile,
                height = 10, width = 8, dpi = 400)
print(paste0("Saving plot to ", plotFile))
```

Finally we save the seasonal per-household profiles for future re-use.

```{r save household profiles}
# save the profiles for future use ----
ofile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_byHouseholdSeasonalProfiles.csv")
print(paste0("Saving profile data used to build this plot to: ", ofile, "..."))
data.table::fwrite(plotDT, ofile)

cmd <- paste0("gzip -f ", "'", path.expand(ofile), "'") # gzip it - use quotes in case of spaces in file name, expand path if needed
try(system(cmd)) # in case it fails - if it does there will just be .csv files (not gzipped) - e.g. under windows
print(paste0("Gzipped ", ofile))

t <- summary(plotDT)

kableExtra::kable(caption = paste0("Summary of household level mean profiles for ", circuitPattern),
             t)
```

As we can see there is considerable variation between households in both the level and timing of heat pump demand.

Note that the code saves a high definition version of the plot and the profiles for future re-use.

## Profile plots: overall household mean

This section shows a plot of mean and median profiles across all household by season. The mean profile also shows the level of variance by plotting error bars at +/- 1 s.d.

Figure \@ref{fig:meanPlot} shows the mean power for circuits containing "`r circuitPattern`" by season.

```{r meanPlot, fig.cap="Mean power profiles (with 95% CI)", fig.height=6}
# create plot table
plotDT <- gs1MinDT[!is.na(season), 
                     .(meanW = mean(powerW),
                       medianW = median(powerW),
                       nObs = .N,
                       sdW = sd(powerW)), keyby = .(obsHourMin, season)]

# mean plot ----
# basic version (mean line only)
myPlot <- ggplot2::ggplot(plotDT, aes(y = meanW, x = obsHourMin, colour = season)) +
  labs(caption = paste0(plotCaption,"\nn households = ", nHHs),
       x = "Time of Day")

labelProfilePlot(myPlot + geom_line())

# save the plot at high def
plotFile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_overallMeanSeasonalProfilePlot.pdf")
ggplot2::ggsave(plotFile,
                height = 10, width = 8, dpi = 400)
print(paste0("Saving plot to ", plotFile))
```

Figure \@ref{fig:meanPlotSD} plot adds +/1 1 standard deviation to show the degree of variation/uncertainty.

> NB: this erroneously assumes a normal distribution

```{r meanPlotSD, fig.cap="Mean power profiles (with +/- 1 SD)", fig.height=6}

# add error bars and so need to facet
labelProfilePlot(myPlot + geom_errorbar(aes(ymin = meanW - sdW, ymax = meanW + sdW, 
                                            colour = season), 
                                        width = 0.4, alpha = 0.2) +
                   facet_grid(season ~ .) +
                   geom_point() +
                   labs( caption = paste0(plotCaption,"\nn households = ", nHHs,
                                                 "\nError bars = +/- 1 S.D."))
                 )

# save the plot at high def
plotFile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_overallMeanSDSeasonalProfilePlot.pdf")
ggplot2::ggsave(plotFile,
                height = 10, width = 8, dpi = 400)
print(paste0("Saving plot to ", plotFile))
```

Figure \@ref{fig:meanPlotCI} plot adds 95% confidence intervals to show the degree of variation/uncertainty. Due to the highly skewed nature of the power demand data, this is implemented using the ggplot2 `mean_cl_boot()` summary function. This is an implementation of [smean.cl.boot](https://www.rdocumentation.org/packages/Hmisc/versions/4.1-1/topics/smean.sd), a "basic nonparametric bootstrap for obtaining confidence limits for the population mean without assuming normality" [source](https://stackoverflow.com/questions/17414565/interpretation-of-stat-summary-mean-cl-boot-at-ggplot2).

```{r meanPlotCI, fig.cap="Mean power profiles (with 95% CI)"}
# 95% CI as a variance indicator
# uses mean_cl_boot() so does not need to assume normality
# slow as uses bootstrap

# add error bars
labelProfilePlot(ggplot(gs1MinDT, aes(x = obsHourMin, 
                                      group = season,
                                      colour = season,
                                      y = powerW)
                        ) + 
                   stat_summary(fun.data = "mean_cl_boot") +
                   labs( caption = paste0(plotCaption,"\nn households = ", nHHs,
                                                 "\nError bars = 95% CI"))
                 )

# save the plot at high def
plotFile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_overallMeanCISeasonalProfilePlot.pdf")
ggplot2::ggsave(plotFile,
                height = 10, width = 8, dpi = 400)
print(paste0("Saving plot to ", plotFile))
```

Figure \@ref{fig:medianPlot} shows the median for the same combinations of data.

```{r medianPlot, fig.cap="Median power plot"}
# median plot ----
myPlot <- ggplot2::ggplot(plotDT, aes(y = medianW, x = obsHourMin, colour = season)) +
  geom_point() +
  facet_grid(season ~ .) +
  labs(caption = paste0(plotCaption,"\nn households = ", nHHs),
       x = "Time of Day")
labelProfilePlot(myPlot)

# save the plot at high def
plotFile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_overallMedianSeasonalProfilePlot.pdf")
ggplot2::ggsave(plotFile,
                height = 10, width = 8, dpi = 400)
print(paste0("Saving plot to ", plotFile))
```

Finally we save the seasonal aggregates for future re-use.

```{r saveSeasonalAggregates}
# save the profiles for future use ----
ofile <- paste0(outPath, "profiles/", circuitPattern, "_", dateFrom, "_", dateTo, "_overallSeasonalProfiles.csv")
print(paste0("Saving profile data used to build this plot to: ", ofile, "..."))
data.table::fwrite(plotDT, ofile)

cmd <- paste0("gzip -f ", "'", path.expand(ofile), "'") # gzip it - use quotes in case of spaces in file name, expand path if needed
try(system(cmd)) # in case it fails - if it does there will just be .csv files (not gzipped) - e.g. under windows
print(paste0("Gzipped ", ofile))

t <- summary(plotDT)

kableExtra::kable(caption = paste0("Summary of overall profiles for ", circuitPattern),
             t)
```

The difference between the mean and median plots is instructive - it suggests that the mean plots for summer are skewed by a few higher heat pump-using households.

The plots could be repeated or re-facted e.g. by household size.

As before, the code saves a high definition version of the plot.

## Profile plots: overall household mean by household size.

Figure \@ref{fig:meanPlot} shows the mean power for circuits containing "`r circuitPattern`" by season and presenceo f children. This uses the linkID to correctly link all household attribute data to the grid spy data.

```{r meanPlotHHSize, fig.cap="Mean power profiles by household size", fig.height=6}
# create plot table

```

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * kableExtra - to create this document & neat tables [@knitr]
 * GREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
