---
params:
  subtitle: ""
  title: ""
  version: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r ggrParams$Authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "ggplot2", # for fancy graphs
             "readr", # for reading & parsing .csv files
             "lubridate" # for year()
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# Local parameters ----

b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- paste0(ggrParams$projLoc, "/docs/plots/") # where to put the plots

  
# Local functions ----


```

\newpage

# About

## Report circulation:

 * Public – this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

```{r citation, child=ggrParams$citation}
```

## History

```{r history, child=ggrParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridData/commits/master/makeDocs/buildOverviewReport.Rmd)
 
## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
This report provides an overview of the GREEN Grid project [@stephenson_smart_2017] research data.

# Data Package

Version `r params$version` of the data package contains:

 * checkPlots: 
    - simple line charts of mean power per month per year for each circuit monitored for each household. These are a useful check;
    - tile plots (heat maps/carpet plots) of the number of observations per hour per day. Also a useful check;
 * powerData: 1 minute power demand data for each household. One file per household;
 * surveyData: anonymised household attribute data.

# Study recruitment

```{r loadSafeHouseholdData, include=FALSE}
hhDT <- data.table::as.data.table(readr::read_csv(ggrParams$hhAttributes))
```

```{r loadHHgsStat, include=FALSE}
gsStatDT <- data.table::as.data.table(readr::read_csv(paste0(ggrParams$statsLoc, "hhStatsByDate.csv")))
```

The project research sample comprises `r nrow(hhDT)` households who were recruited via the local power lines companies in two areas: New Plymouth starting in May 2014 and Hawkes Bay starting in November 2014. 

Recruitment was via a non-random sampling method and a number of households were intentionally selected for their 'complex' electricity consumption (and embedded generation) patterns and appliances [@dianaThesis2015, @stephenson_smart_2017, @jack_minimal_2018, @suomalainen_comparative_2017]. 

The lines companies invited their own employees and those of other local companies to participate in the research and ~80 interested potential participants completed short or long forms of the Energy Cultures 2 household survey [@ec2Survey2015]. Households were then selected from this pool by the project team based on selection criteria relevant to the GREEN Grid project. These included:

 * having the majority of their energy supply from electricity (i.e. not gas heating);
 * household size;
 * types of appliances owned.

After informed consent was obtained from each household, an electrician contracted by the two lines companies completed an appliance survey to record detailed information about the appliances in each house. This survey contained information about the number of appliances owned, brand, model number, efficiency and age. The electrician also installed the GridSpy units which recorded electricity power demand at a circuit level. The GridSpy units automatically upload the monitoring data to the GridSpy company's secure database from where it was downloaded by the GREEN Grid research team. 

As a result of this process the sample cannot be assumed to represent the population of customers (or employees) of any of the companies involved, nor the populations in each location [@stephenson_smart_2017].

Table \@ref(tab:sampleTable) shows the number in each sample.

```{r sampleTable}
t <- hhDT[, .(nHouseholds = .N), keyby = .(Location)]

kableExtra::kable(t, caption = "Sample location") %>%
  kable_styling()
```

Table \@ref(tab:sampleSurveys) shows the number for whom valid appliance and survey data is available in this data package. Note that even those which appear to lack appliance data may have sufficient survey data to deduce appliance ownership (see question numbers `Q19_*` and `Q40_*`).

```{r sampleSurveys}

t <- hhDT[, .(nHouseholds = .N), keyby = .(Location, hasShortSurvey, hasLongSurvey, hasApplianceSummary)]

kableExtra::kable(t, caption = "Sample information") %>%
  kable_styling()
```

# Data collection duration

Figure \@ref(fig:liveDataHouseholds) shows the total number of households for whom GridSpy data exists on a given date by sample. The plot includes _any_ data, including partial data and suggests that for analytic purposes the period from April 2015 to March 2016 (indicated) would offer the maximum number of households.

```{r liveDataHouseholds, fig.cap="Number of households sending GridSpy data by date"}
setkey(gsStatDT, linkID)
dt <- merge(gsStatDT, hhDT, allow.cartesian=TRUE)

plotDT <- dt[, .(nHH = uniqueN(hhID)), keyby = .(date, Location)]

# point plot ----
myCaption <- paste0("Source: GREEN Grid Project data ", min(dt$date), " to ", max(dt$date))
rectAlpha <- 0.3
vLineAlpha <- 0.8
vLineCol <- "#0072B2" # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette

yMin <- min(plotDT$nHH)
yMax <- 2 * max(plotDT$nHH)

ggplot2::ggplot(plotDT, aes( x = date, y = nHH, fill = Location)) +
  geom_col() +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 0.5)) + 
  labs(title = "N live households per day for all clean GridSpy data",
       caption = paste0(myCaption, "\nShaded area indicates 12 month period with largest number of households"),
       y = "Number of households"
       
  ) + 
  annotate("rect", xmin = as.Date("2015-04-01"),
             xmax = as.Date("2016-04-01"), 
             ymin = yMax, ymax = yMin, alpha = rectAlpha, fill = vLineCol)
```

# Key attributes

Table \@ref(tab:allHhData) shows key attributes for the recruited sample. Note that two GridSpy monitors were re-used and so require new hhIDs to be set from the dates given. This has already been done in the [clean GridSpy data](gridSpy1mProcessingReport.html#41_recoding_re-allocated_grid_spy_units) for the relevant households. Linkage between the survey and GridSpy data should always use `linkID` to avoid errors.

```{r allHhData}
kableExtra::kable(hhDT[, c("hhID", "linkID", "Location", "notes", "r_stopDate")][order(linkID)], 
             caption = "Sample details") %>%
  kable_styling()

```

# Known issues

We maintain a [known data issues list](https://github.com/CfSOtago/GREENGridData/labels/dataIssue) via our GitHub repository. If you think there is a data issue please check the repo list first and then add a new one if appropriate.

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * knitr - to create this document & neat tables [@knitr]
 * GREENGridData - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
