---
params:
 title: ""
 subtitle: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r ggrParams$Authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace # <- don't do this - rmeoves params set in yaml!

# Set start time ----
startTime <- proc.time()

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # reading/writing csv files
             "readxl", # reading xlsx
             "skimr", # for skim
             "knitr" # for kable & captions
)
# load them
nzGREENGridDataR::loadLibraries(rmdLibs)


# Local paramaters
  
```

\newpage

# About

## Report circulation:

 * Public – this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * `r ggrParams$Authors` (`r lubridate::year(today())`) `r params$title` `r params$subtitle`, `r ggrParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

```{r includeHistory, child=ggrParams$historyGenericRmd}
```
 
Specific history of this report:

 * https://github.com/dataknut/nzGREENGridDataR/commits/master/dataProcessing/surveys/createHouseholdAttributes.Rmd

## Requirements:

This report uses the original NZ GREEN Grid household survey and meta data.

## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
The purpose of this report is to: 

 * create a household attribute file that can be linked to the project power monitoring data;
 * describe the household attribute data.

The resulting cleaned data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners.

The data contains a unique household id (`linkID`) which can be used to link it to the NZ GREEN Grid time use diaries and dwelling/appliance surveys.

# Load data

## Household master dataset
Table \@ref(tab:loadMasterData) shows the number of households in each area. 

```{r loadMasterData}
hhMasterDT <- nzGREENGridDataR::getHouseholdData(ggrParams$gsHHMasterFile) # cleans it as it loads
setkey(hhMasterDT, linkID)
t <- with(hhMasterDT, table(Location, useNA = "always"))
knitr::kable(caption = "Sample location (master file)", t)
```

In total we have `r nrow(hhMasterDT)` households in two sample areas.

## Appliance data

Household appliance ownership was recorded during recruitment using a detailed survey. However this data is not readily available for all households at present. Table \@ref(tab:applianceTable) shows the number of households in each area for whom summary appliance data exists. 

```{r loadApplianceData}
hhAppliancesDT <- data.table::as.data.table(readxl::read_xlsx(ggrParams$ApplianceData))
```

```{r applianceTable}
hhAppliancesDT$sample <- NULL # too detailed
setkey(hhAppliancesDT, linkID)
t <- with(hhAppliancesDT[hhMasterDT], table(Location, hasApplianceSummary, useNA = "always"))
knitr::kable(caption = "Sample (appliance summary file, NA indicates no data)", t)
```

The appliances recorded in this summary are shown in Table \@ref(tab:applianceNames).

```{r applianceNames}
xcols <- c("linkID", "hasApplianceSummary")
al <- names(hhAppliancesDT[, -..xcols])
knitr::kable(caption = "Appliances recorded", al)
```

## Household survey data

The Energy Cultures 2 survey [@ec2Survey2015] was used to collect data on household energy and transport attitudes/behaviours and usages for one of the regional samples only. The long form of the survey was used for some households and the short form for others.

```{r loadSafeSurvey, include=FALSE}
hhAttributesDT <- data.table::as.data.table(readr::read_csv(ggrParams$hhAttributes))
```

```{r responseTable}
t <- hhAttributesDT[, .(nHouseholds = .N), keyby = .(Location, hasLongSurvey, hasShortSurvey)]
knitr::kable(t, caption = "Survey responses (NA indicates no survey)")
```

# Final household attribute data description

Table \@ref(tab:skimTable) reports the final household attribute variables and their internal distributions. The question labels are found in Table \@ref(tab:loadEc2Labels) below.

```{r summariseHhAttributes, include=FALSE}
#Hmisc::describe(hhAttributesDT)
skimDF <- skimr::skim(hhAttributesDT)
```

```{r skimTable, results="asis"}
#https://github.com/ropensci/skimr/blob/master/README.md

skimr::kable(skimDF, caption = "Descriptive summary (all variables)")
```


# Household attribute frequency tables

This section reports tables of the key household attributes by sample location. Note that NA usually means not known.

## Number of adults

```{r nAdults}
t <- with(hhAttributesDT, table(nAdults, Location, useNA = "always"))
knitr::kable(caption = "Number of adults in household by location", t)
```

## Number of teenagers

```{r nTeens}
t <- with(hhAttributesDT, table(nTeenagers13_18, Location, useNA = "always"))
knitr::kable(caption = "Number of teenagers in household by location", t)
```

## Number of children

```{r nChildren}
t <- with(hhAttributesDT, table(nChildren0_12, Location, useNA = "always"))
knitr::kable(caption = "Number of children in household by location", t)
```

## Notes variable

The `notes` column may have been set for any number of reasons and measn the monitoring data for a given household should be used with caution.

```{r outliers}
t <- with(hhAttributesDT, table(notes, Location, useNA = "always"))
knitr::kable(caption = "Notes by location", t)
```

# Summary

The following table (\@ref(tab:dataHeader)) shows the key columns of the household attributes file. The data can be linked to the gridSpy data using `linkID`. `linkID` is used to flag the two [re-used GridSpy units](gridSpy1mProcessingReport.html#41_recoding_re-allocated_grid_spy_units). As an example, unit rf_15 was re-used in a different household. We have therefore created `linkID` so that the correct household data (rf_15a or rf_15b) can be linked to the GridSpy data (coded rf_15) at the correct date.

Note also that data exists in the GridSpy power demand data for whom no household data exists (e.g. rf_01 & rf_02).

```{r dataHeader}
knitr::kable(caption = "Household data: key columns", 
             hhAttributesDT[, c("hhID", "linkID", "Location", "nAdults", "r_stopDate", "notes")][order(hhID)])
```

# Runtime

```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for csv reading/writing [@readr]
 * skimr - for skim [@skimr]
 * knitr - for kable [@knitr]
 * readxl - reading .xlsx [@readxl]
 * nzGREENGridDataR - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# Energy Cultures 2 Long Survey Questions
Table \@ref(tab:loadEc2Labels) lists the full Energy Cultures 2 survey [@ec2Survey2015]. Only a subset of these variables are currently available in the GREEN Grid safe data package. Requests to add further variables should be made via a [GitHub repository issue](https://github.com/dataknut/nzGREENGridDataR/labels/variableRequest). 

```{r loadEc2Labels}
ec2LabsDF <- readxl::read_xlsx(ggrParams$ec2LongSurveyLabels, sheet = "ec2Long")

knitr::kable(caption = "Energy Cultures 2 Survey labels (long)", ec2LabsDF)
```

# References
