---
params:
 title: ""
 subtitle: ""
 version: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r ggrParams$Authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
bibliography: '`r paste0(GREENGridData::findParentDirectory("GREENGridData"), "/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace # <- don't do this - rmeoves params set in yaml!

# Set start time ----
startTime <- proc.time()

# Packages needed in this .Rmd file ----
rmdLibs <- c("readxl", # reading xlsx
             "skimr" # for skim
)
# load them
GREENGridData::loadLibraries(rmdLibs)


# Local paramaters
  
```

\newpage

# About

## Report circulation:

 * Public – this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

If you wish to use any of the material from this report please cite as:

 * `r ggrParams$Authors` (`r lubridate::year(today())`) `r params$title` `r params$subtitle`, `r ggrParams$pubLoc`.

This work is (c) `r lubridate::year(today())` the University of Southampton.

## History

```{r history, child=ggrParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridData/commits/master/makeDocs/buildHouseholdAttributesReport.Rmd)

## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
The purpose of this report is to: 

 * create a household attribute file that can be linked to the project power monitoring data;
 * describe the household attribute data.

The resulting cleaned data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners.

The data contains a unique household id (`linkID`) which can be used to link it to the NZ GREEN Grid time use diaries and dwelling/appliance surveys.

# Load data

## Household master dataset
Table \@ref(tab:loadMasterData) shows the number of households in each area. 

```{r loadMasterData}
hhMasterDT <- GREENGridData::getHouseholdData(ggrParams$gsHHMasterFile) # cleans it as it loads
setkey(hhMasterDT, linkID)
t <- with(hhMasterDT, table(Location, useNA = "always"))
kableExtra::kable(caption = "Sample location (master file)", t) %>% 
  kable_styling()
```

In total we have `r nrow(hhMasterDT)` households in two sample areas.

## Appliance data

Household appliance ownership was recorded during recruitment using a detailed survey. However this data is not readily available for all households at present. Table \@ref(tab:applianceTable) shows the number of households in each area for whom summary appliance data exists. 

```{r loadApplianceData}
hhAppliancesDT <- data.table::as.data.table(readxl::read_xlsx(ggrParams$ApplianceData))
```

```{r applianceTable}
hhAppliancesDT$sample <- NULL # too detailed
setkey(hhAppliancesDT, linkID)
t <- with(hhAppliancesDT[hhMasterDT], table(Location, hasApplianceSummary, useNA = "always"))
kableExtra::kable(caption = "Sample (appliance summary file, NA indicates no data)", t) %>%
  kable_styling()
```

The appliances recorded in this summary are shown in Table \@ref(tab:applianceNames).

```{r applianceNames}
xcols <- c("linkID", "hasApplianceSummary")
al <- names(hhAppliancesDT[, -..xcols])
kableExtra::kable(caption = "Appliances recorded", al) %>%
  kable_styling()
```

## Household survey data

The Energy Cultures 2 survey [@ec2Survey2015] was used to collect data on household energy and transport attitudes/behaviours and usages for one of the regional samples only. The long form of the survey was used for some households and the short form for others.

```{r loadSafeSurvey, include=FALSE}
# read in data produced by processHouseholdAttributes.R
# do not do any re-coding here - put that in the above .R script
hhAttributesDT <- data.table::as.data.table(readr::read_csv(ggrParams$hhAttributes))
```

```{r responseTable}
t <- hhAttributesDT[, .(`n Households` = .N), keyby = .(Location, hasLongSurvey, hasShortSurvey)]
kableExtra::kable(t, caption = "Survey responses (NA indicates no survey)") %>%
  kable_styling()
```

# Final household attribute data description

Table \@ref(tab:skimTable) reports the final household attribute variables and their internal distributions. The question labels are found in Table \@ref(tab:loadEc2Labels) below.

```{r summariseHhAttributes, include=FALSE}
#Hmisc::describe(hhAttributesDT)
skimDF <- skimr::skim(hhAttributesDT)
```

```{r skimTable, results="asis"}
#https://github.com/ropensci/skimr/blob/master/README.md

skimr::kable(skimDF, caption = "Descriptive summary (all variables)")
```


# Household attribute frequency tables

This section reports tables of the key household attributes by sample location. Note that NA usually means not known.

## Number of adults

```{r nAdults}
t <- with(hhAttributesDT, table(nAdults, Location, useNA = "always"))
kableExtra::kable(caption = "Number of adults in household by location", t) %>%
  kable_styling()
```

## Number of teenagers

```{r nTeens}
t <- with(hhAttributesDT, table(nTeenagers13_18, Location, useNA = "always"))
kableExtra::kable(caption = "Number of teenagers in household by location", t) %>%
  kable_styling()
```

## Number of children

```{r nChildren}
t <- with(hhAttributesDT, table(nChildren0_12, Location, useNA = "always"))
kableExtra::kable(caption = "Number of children in household by location", t) %>%
  kable_styling()
```

## Notes variable

The `notes` column may have been set for any number of reasons and measn the monitoring data for a given household should be used with caution.

```{r outliers}
t <- with(hhAttributesDT, table(notes, Location, useNA = "always"))
kableExtra::kable(caption = "Notes by location", t) %>%
  kable_styling()
```

# Summary

The following table (\@ref(tab:dataHeader)) shows the key columns of the household attributes file. The data can be linked to the gridSpy data using `linkID`. `linkID` is used to flag the two [re-used GridSpy units](gridSpy1mProcessingReport.html#41_recoding_re-allocated_grid_spy_units). As an example, unit rf_15 was re-used in a different household. We have therefore created `linkID` so that the correct household data (rf_15a or rf_15b) can be linked to the GridSpy data (coded rf_15) at the correct date.

Note also that data exists in the GridSpy power demand data for whom no household data exists (e.g. rf_01 & rf_02).

```{r createTable, echo=FALSE, include=FALSE}

dt <- hhAttributesDT[, c("hhID", "linkID", "Location", "nAdults", "nChildren0_12", "r_stopDate",
                                "Electric heater", "Heat pump number", "Other Appliance", "PV Inverter", "Energy Storage","Other Generation Device",
                                "notes")]
# clean up table
for (j in names(dt)) {
  set(dt,which(is.na(dt[[j]])),j,".")
}
```

```{r dataHeader}
kableExtra::kable(caption = "Household data: key columns (NA set to '.' in string variables for clarity)", 
             dt[order(hhID)]) %>%
  kable_styling()
```

# Runtime

```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * GREENGridData [@GREENGridData] which depends on:
    - data.table [@data.table]
    - dplyr [@dplyr]
    - hms [@hms]
    - lubridate [@lubridate]
    - progress [@progress]
    - readr [@readr]
    - readxl [@readxl]
    - reshape2 [@reshape2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * rmarkdown [@rmarkdown]
 * skimr [@skimr]
 
## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# Energy Cultures 2 Long Survey Questions
Table \@ref(tab:loadEc2Labels) lists the full Energy Cultures 2 survey [@ec2Survey2015]. Only a subset of these variables are currently available in the GREEN Grid safe data package. Requests to add further variables should be made via a [GitHub repository issue](https://github.com/CfSOtago/GREENGridData/labels/variableRequest). 

```{r loadEc2Labels}
ec2LabsDF <- readr::read_csv(ggrParams$ec2LongSurveyLabels)

kableExtra::kable(caption = "GREEN Grid/Energy Cultures 2 Survey labels (long)", ec2LabsDF) %>%
  kable_styling()
```

# References
