---
params:
  subtitle: ""
  title: ""
  version: ""
  localData: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r ggrParams$Authors`'
date: 'Last run at: `r getRunDateTime()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
bibliography: '`r paste0(GREENGridData::findParentDirectory("GREENGridData"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()

# Libraries
library(GREENGridData)
GREENGridData::setup()

localLibs <- c("data.table", "ggplot2", "kableExtra")
GREENGridData::loadLibraries(localLibs)
# Local parameters ----

b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- paste0(ggrParams$repoLoc, "/docs/plots/") # where to put the plots


# Local functions ----


```

\newpage

# About

## Report circulation:

 * Public – this report is intended to accompany the data release.
 
## License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
## Citation

```{r citation, child=ggrParams$citation}
```

## History

```{r history, child=ggrParams$history}
```
 * this [report's edit history](https://github.com/CfSOtago/GREENGridData/commits/master/makeDocs/checkNegativePower.Rmd)
 
## Support

```{r generic support, child=ggrParams$supportGeneric}
```
 
\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 
This report provides an analysis of negative power values to be found in the GREEN Grid project [@stephenson_smart_2017] research data.

There are a number of observations that have recorded negative power. There are at least two potential reasons for this:

 * a Grid Spy sensor was placed on the wrong wire and/or set to the wrong phase
 * a circuit contained a PV inverter which pushes power into the home and which is measured as -ve power (when active)
 
The following analysis of the incidence of negative values makes recommendations on actions to take. 

```{r loadSafeHouseholdData, include=FALSE}
hhDT <- data.table::as.data.table(readr::read_csv(ggrParams$hhAttributes))
```

The project research sample comprises `r nrow(hhDT)` households. Table \@ref(tab:sampleSurveys) shows the number for whom valid appliance and survey data is available in this data package. Note that even those which appear to lack appliance data may have sufficient survey data to deduce appliance ownership (see question numbers `Q19_*` and `Q40_*`).

```{r sampleSurveys}

t <- hhDT[, .(nHouseholds = .N), keyby = .(Location, hasShortSurvey, hasLongSurvey, hasApplianceSummary)]

kableExtra::kable(t, caption = "Sample information") %>%
  kable_styling()
```

# Key attributes

Table \@ref(tab:allHhData) shows key attributes for the recruited sample. Note that two GridSpy monitors were re-used and so require new hhIDs to be set  from the date of re-use using the `linkID` variable. This is explained in more detail in the [GridSpy processing](gridSpy1mProcessingReport_v1.0.html#reallocation) report. Linkage between the survey and GridSpy data should therefore always use `linkID` to avoid errors.

```{r allHhData}
keyCols <- c("hhID", "linkID", "Location", "surveyStartDate", "nAdults",
             "notes", "r_stopDate", "hasApplianceSummary",
             "PV Inverter", "Energy Storage")
kableExtra::kable(hhDT[, ..keyCols ][order(linkID)], 
             caption = "Sample details") %>%
  kable_styling()

nPV <- table(hhDT$`PV Inverter`)[[1]]
```

As we can see there were `r nPV` households with PV inverters. These were:

```{r listPV}
kableExtra::kable(hhDT[`PV Inverter` == "yes", ..keyCols ][order(linkID)], 
             caption = "Sample details (PV households)") %>%
  kable_styling()

# make easy to use flag
hhDT <- hhDT[, hasPV := "No PV" ]
hhDT <- hhDT[`PV Inverter` == "yes", hasPV := "Has PV" ]
```


# Check negative power

We would rather not load the entire data for all households to do this so instead we start by investigating negative power in extracts for three different circuit labels:

 * Heat Pump
 * Hot Water
 * Lighting
 
These were extracted from the processed data using the [example extraction code](https://github.com/CfSOtago/GREENGridData/blob/master/examples/code/extractCleanGridSpy1minCircuit.R) provided in this package. 

## Negative power: Heat Pump

As Figure \@ref(fig:hpBoxplot) shows negative power observations for Heat Pumps are spread across a number of households. It also shows that there is at least one very large outlier reading for rf_27.

```{r loadHeatPump}
dataExtracts <- paste0(ggrParams$dataLoc, "cleanData/safe/gridSpy/1min/dataExtracts/")
hpDT <- GREENGridData::getCleanGridSpyFile(paste0(dataExtracts, "Heat Pump_2015-04-01_2016-03-31_observations.csv.gz")) # use this function to ensure proper number etc formats
summary(hpDT)

```

```{r hpBoxplot, fig.cap="Boxplot of Heat Pump power values per household"}
hpBox <- ggplot2::ggplot(hpDT, aes(x = linkID, y = powerW, group = linkID)) +
  geom_boxplot()

hpBox + coord_flip()
```

As we can see from Table \@ref(tab:maxOutlier) there is only 1 value where power > 10kW. We recommend that this value be excluded from analysis.

```{r maxOutlier}
t <- head(hpDT[powerW > 10000])

kableExtra::kable(t, caption = "All Heat Pump records where power > 10kW")
```


```{r hpNeg, fig.cap="Number of negative power observations per hour (max = 60 * 60) by household over time"}
# aggregate to days
hourlyHpDT <- hpDT[, .(minPower = min(powerW), 
                       maxPower = max(powerW),
                       nObs = .N), 
                   keyby = .(linkID, obsHourDate = lubridate::floor_date(r_dateTime, unit = "hour"))]

# match to hhDT
setkey(hourlyHpDT, linkID)
setkey(hhDT, linkID)
keepCols <- c("linkID", "hasPV")
#hourlyHpDT <- hhDT[, ..keepCols][hourlyHpDT]

hpNegTile <- ggplot2::ggplot(hourlyHpDT[minPower < 0], aes(x = obsHourDate, y = linkID, fill = nObs)) +
  geom_tile()

hpNegTile + scale_fill_gradient(low = "green", high = "red") +
  labs(x = "Hour of day")
```

Turning to the negative values, Figure \@ref(fig:hpNeg) shows that the vast majority of negative power values are due to one household (rf_46, which does not have PV). The others appear on isolated days and are most likely to be due to instrument installation error. 

Table \@ref(tab:negHpTable) lists the households where any negative heat pump power values are observed while Figure \@ref(fig:negHpTable) shows a density plot for each household. The latter (together with Figure \@ref(fig:hpNeg)) suggests that the negative values continually observed for rf_46 may constitute some form of intermittent error since positive values were also recorded (c.f. Table \@ref(tab:negHpTable)). The other households on the other hand report a few negative observations suggesting instrument installation error that was quickly corrected.

```{r negHpTable, fig.cap="Histogram of negative power values only"}
negHHsHpDT <- hpDT[, .(minPower = min(powerW),
                   meanPower = mean(powerW),
                   maxPower = max(powerW)
                   ),
               keyby = .(linkID)]

kableExtra::kable(negHHsHpDT[minPower < 0], caption = "Households where negative power observed")

negHH <- negHHsHpDT[minPower < 0, .(linkID, minPowerHP = minPower, maxPowerHP = maxPower)]

setkey(negHH, linkID)

ggplot(hpDT[powerW < 0], aes(x = powerW, fill = linkID)) + geom_histogram() + facet_grid(linkID ~ ., scales = "free_y") + 
  labs(caption = "Note y scale varies across panel")
```


We therefore recommend that when analysing heat pump data:

 * rf_46 is removed;
 * this value is removed: `rf_27 	2015-08-22 10:33:00 	Heat Pump$2826 	27759 `
 * _any_ day with negative power values for any household is also removed.
 
## Negative power: Hot Water

As Figure \@ref(fig:hwBoxplot) shows negative power observations for Hot Water are spread across a number of households. In this case though, there are no very large positive outlier readings.

```{r loadHotWater}

hwDT <- GREENGridData::getCleanGridSpyFile(paste0(dataExtracts, "Hot Water_2015-04-01_2016-03-31_observations.csv.gz")) # use this function to ensure proper number etc formats
summary(hwDT)

```

```{r hwBoxplot, fig.cap="Boxplot of Hot Water power values per household"}
hwBox <- ggplot2::ggplot(hwDT, aes(x = linkID, y = powerW, group = linkID)) +
  geom_boxplot()

hwBox + coord_flip()
```


```{r hwNeg, fig.cap="Number of negative power observations per hour (max = 60 * 60) by household over time"}
# aggregate to days
hourlyHwDT <- hwDT[, .(minPower = min(powerW), 
                       maxPower = max(powerW),
                       nObs = .N), 
                   keyby = .(linkID, obsHourDate = lubridate::floor_date(r_dateTime, unit = "hour"))]

# match to hhDT
setkey(hourlyHpDT, linkID)
setkey(hhDT, linkID)
keepCols <- c("linkID", "hasPV")
#hourlyHpDT <- hhDT[, ..keepCols][hourlyHwDT]

hwNegTile <- ggplot2::ggplot(hourlyHwDT[minPower < 0], aes(x = obsHourDate, y = linkID, fill = nObs)) +
  geom_tile()

hwNegTile + scale_fill_gradient(low = "green", high = "red") +
  labs(x = "Hour of day")
```

Turning to the negative values, Figure \@ref(fig:hwNeg) shows that the vast majority of negative power values are again due to one household (rf_46, which does not have PV) but with the addition of a number from rf_14. The others appear on isolated days and as before are most likely to be due to instrument installation error. 

Table \@ref(tab:negHwTable) lists the households where any negative hot water power values are observed while Figure \@ref(fig:negHwTable) shows a density plot for each household. Note that these households do not necessarily match to those reported above for heat pumps partly because not all households have heat pumps. The latter (together with Figure \@ref(fig:hwNeg)) suggests that the negative values continually observed for rf_46 may constitute some form of intermittent error since positive values were also recorded (c.f. Table \@ref(tab:negHwTable)). The other households on the other hand (including rf_14) report negative observations on specific days suggesting instrument installation error that was quickly corrected.

```{r negHwTable, fig.cap="Histogram of negative power values only"}
negHHsHwDT <- hwDT[, .(minPower = min(powerW),
                   meanPower = mean(powerW),
                   maxPower = max(powerW)
                   ),
               keyby = .(linkID)]

kableExtra::kable(negHHsHwDT[minPower < 0], caption = "Households where negative power observed")
dt <- negHHsHwDT[minPower < 0, .(linkID, minPowerHW = minPower, maxPowerHW = maxPower)]

negHH <- merge(negHH, dt, all = TRUE)

ggplot(hwDT[powerW < 0], aes(x = powerW, fill = linkID)) + geom_histogram(binwidth = 10) + facet_grid(linkID ~ ., scales = "free_y") + 
  labs(caption = "Note y scale varies across panel")
```

We therefore recommend that when analysing heat pump data:

 * rf_46 is removed;
 * _any_ day with negative power values is also removed (including for rf_14).

## Lighting

As Figure \@ref(fig:lBoxplot) shows negative power observations for Lighting are found in only one household. There are also some relatively large values across at least 2 or 3 households which may indicate that the lighting circuits did not just power lights.

```{r loadLightingWater}

lDT <- GREENGridData::getCleanGridSpyFile(paste0(dataExtracts, "Lighting_2015-04-01_2016-03-31_observations.csv.gz")) # use this function to ensure proper number etc formats
summary(hwDT)

```

```{r lBoxplot, fig.cap="Boxplot of Hot Water power values per household"}
lBox <- ggplot2::ggplot(lDT, aes(x = linkID, y = powerW, group = linkID)) +
  geom_boxplot()

lBox + coord_flip()
```

This is clear when we consider the distribution of power by household ID and circuit name (Table \@ref(tab:lCheck)). Analysis of 'lighting' for these households should therefore be done with care.

```{r lCheck}
t <- lDT[, .(minPower = min(powerW),
             maxPower = max(powerW)),
         keyby = .(linkID, circuit)]
kableExtra::kable(t[order(-maxPower)], caption = "Power distribution by household and circuit label")
```

```{r lNeg, fig.cap="Number of negative power observations per hour (max = 60 * 60) by household over time"}
# aggregate to days
hourlyDT <- lDT[, .(minPower = min(powerW), 
                       maxPower = max(powerW),
                       nObs = .N), 
                   keyby = .(linkID, obsHourDate = lubridate::floor_date(r_dateTime, unit = "hour"))]

# match to hhDT
setkey(hourlyDT, linkID)
setkey(hhDT, linkID)
keepCols <- c("linkID", "hasPV")
#hourlyHpDT <- hhDT[, ..keepCols][hourlyHwDT]

lNegTile <- ggplot2::ggplot(hourlyDT[minPower < 0], aes(x = obsHourDate, y = linkID, fill = nObs)) +
  geom_tile()

lNegTile + scale_fill_gradient(low = "green", high = "red") +
  labs(x = "Hour of day")
```

Turning to the negative values, Figure \@ref(fig:lNeg) shows that all negative power values are again due to one household (rf_46, which does not have PV). 

Table \@ref(tab:neglTable) lists the households where any negative lighting power values are observed while Figure \@ref(fig:neglTable) shows a density plot for each household. 
The latter (together with Figure \@ref(fig:lNeg)) again suggests that the negative values continually observed for rf_46 may constitute some form of intermittent error since positive values were also recorded (c.f. Table \@ref(tab:neglTable)).

```{r neglTable, fig.cap="Histogram of negative power values only"}
negHHslDT <- lDT[, .(minPower = min(powerW),
                   meanPower = mean(powerW),
                   maxPower = max(powerW)
                   ),
               keyby = .(linkID)]

kableExtra::kable(negHHslDT[minPower < 0], caption = "Households where negative power observed")
dt <- negHHslDT[minPower < 0, .(linkID, minPowerL = minPower, maxPowerL = maxPower)]

negHH <- merge(negHH, dt, all = TRUE)

ggplot(lDT[powerW < 0], aes(x = powerW, fill = linkID)) + geom_histogram(binwidth = 10) + facet_grid(linkID ~ ., scales = "free_y") + 
  labs(caption = "Note y scale varies across panel")
```

We therefore recommend that when analysing heat pump data:

 * rf_46 is removed
 
## Overall patterns of negative observations

Table \@ref(tab:allNeg) shows that there is only one household where negative values are recorded on multiple non-PV circuits. This confirms our view that data for rf_46 should be removed from analysis completely but that other data should be removed on a day by day/household by household basis.

```{r allNeg}
kableExtra::kable(negHH, caption = "All households where negative power observed for heat pumps, hot water or lighting")
```

 
# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * GREENGridData [@GREENGridData] which depends on:
    - data.table [@data.table]
    - dplyr [@dplyr]
    - hms [@hms]
    - lubridate [@lubridate]
    - progress [@progress]
    - readr [@readr]
    - readxl [@readxl]
    - reshape2 [@reshape2]
 * ggplot2 [@ggplot2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
