---
params:
 title: "Economic value of Demand Response"
 subtitle: "Based on Wholesahle Prices 16-17"
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: 'Carsten Dortans (xxx@otago.ac.nz)'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    toc: true
    toc_float: TRUE
    toc_depth: 2
    keep_md: TRUE
    self_contained: no
  bookdown::pdf_document2:
    toc: true
    toc_depth: 2
bibliography: '`r paste0(findParentDirectory("GREENGrid"), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # by default turn off code echo
#options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```

```{r codeSetup, include=FALSE}
#rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- proc.time()

library(GREENGrid)

# Packages needed in this .Rmd file ----
rmdLibs <- c("data.table", # data munching
             "ggplot2", # for fancy graphs
             "lubridate", # date & time processing
             "readr", # writing to files
             "hms", # manipulate h:m:s
             "skimr", # for skim
             "knitr" # for kable
)
# load them
loadLibraries(rmdLibs)


# run gg set up
GREENGrid::setup() # sets a load of parameters which are then accessible via ggParams$xxx

```


# Citation

If you wish to use any of the material from this report please cite as:

 * Dortans, C. (`r 1900 + as.POSIXlt(Sys.Date())$year`) `r params$title`: `r params$subtitle`, `r ggParams$pubLoc`.

This work is (c) `r as.POSIXlt(Sys.time())$year + 1900` the University of Southampton.

\newpage

# About

## Circulation


 
## Purpose

This report is intended to: 

 * load and test GREEN Grid heat pump and hot water profiles.

## Requirements:

 * test dataset stored at `r ggParams$dataLoc`

## History


 
Specific history of this code:

 * https://github.com/CfSOtago/GREENGrid/commits/master/analysis/demandResponse/e



```{r Loading wholesale prices}

WholesalePrices  <- "/Volumes/hum-csafe/Research Projects/GREEN Grid/_RAW DATA/EA_Wholesale_Prices/Wholesale_16-17_clean.csv"


print(paste0("Trying to load: ", WholesalePrices))

```
# Economic analysis of Demand Response
## Reading original data
```{r Calculations}
PricesDT <- data.table::as.data.table(readr::read_csv(WholesalePrices)) # reading data
```

## Time adjustments
```{r Time adjustments}
PricesDT <- PricesDT[, dateTimeStartUTC := lubridate::dmy_hm(`Period start`)] # creating column based on orig. data
PricesDT <- PricesDT[, dateTimeStart := lubridate::force_tz(dateTimeStartUTC, tz = "Pacific/Auckland")] # changing time to NZST
PricesDT$dateTimeStartUTC <- NULL
PricesDT <- PricesDT[, dstFlag := lubridate::dst(dateTimeStart)]
#PricesDT[, .(n = .N), keyby = .(month = lubridate::month(dateTimeStart), dstFlag)]# Daylight saving test
```

## Defining seasons
```{r Defining seasons}

PricesDT <- PricesDT[, month := lubridate::month(dateTimeStart)]

PricesDT <- PricesDT[month >= 9 & month <= 11, season := "Spring"]
PricesDT <- PricesDT[month == 12 | month == 1 | month == 2, season := "Summer"]
PricesDT <- PricesDT[month == 3 | month == 4 | month == 5, season := "Autumn"]
PricesDT <- PricesDT[month == 6 | month == 7 | month == 8, season := "Winter"]

```

## Mean Price per MWh
```{r Mean Price by season, region and half hour}
PricesDT <- PricesDT[, obsHalfHour := hms::as.hms(dateTimeStart)] # creating time column without date


SeasonAvgDT <- PricesDT[, .(meanprice = mean(`Price ($/MWh)`)), keyby = .(season, obsHalfHour, Region)]

```

## Visualisation
```{r test plot}
SeasonAvgDT$season <- factor(SeasonAvgDT$season, levels = c("Spring","Summer",
                                                    "Autumn", "Winter"))

myPlot <- ggplot2::ggplot(SeasonAvgDT, aes(x = obsHalfHour)) +
  geom_line(aes(y=meanprice, color= Region), size=0.5) +
  theme(text = element_text(family = "Cambria")) +
  ggtitle("Test") +
  facet_grid(season ~ .) +
  labs(x='Time of Day', y='$/MWh') +
  scale_x_time(breaks = c(hms::as.hms("00:00:00"), hms::as.hms("03:00:00"), hms::as.hms("06:00:00"),
                          hms::as.hms("09:00:00"), hms::as.hms("12:00:00"),
                          hms::as.hms("15:00:00"), hms::as.hms("18:00:00"), hms::as.hms("21:00:00")))
myPlot


```
# Calculations
## Merging data SC3
```{r merging sc3data}

setkey(SeasonAvgDT, season, obsHalfHour)
sc3dataDT <- as.data.table(sc3data)
setkey(sc3dataDT, season, obsHalfHour)

MergedDT <- sc3dataDT[SeasonAvgDT]


```

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * skimr - for skim [@skimr]
 * knitr - to create this document & neat tables [@knitr]
 * GREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
