---
params:
  localData: ""
title: 'Processing report: NZ GREEN Grid project 1 minute electricity
  power data'
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    code_folding: hide
    fig_caption: true
    keep_md: true
    number_sections: true
    self_contained: no
    toc: true
    toc_float: true
    toc_depth: 2
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
options(knitr.table.format = 'markdown') # try to fix the tables issue (seems to be pushing html into latex)
```


```{r codeSetup, include=FALSE}

# Set start time ----
startTime <- proc.time()

# Local parameters ----

b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576

# Load nzGREENGrid package ----
library(nzGREENGridDataR) # local utilities

nzGREENGridDataR::setup()

# Set grid spy data paths etc from file ----
source(paste0(ggrParams$projLoc, "/dataProcessing/gridSpy/gSpyParams.R"))

# Packages needed in this .Rmd file ----
rmdLibs <- c("ggplot2", # for fancy graphs
             "readr", # for reading & parsing .csv files
             "knitr", # for kable
             "kableExtra" # for extra kable
)
# load them
nzGREENGrid::loadLibraries(rmdLibs)

# Local functions ----


```

\newpage

# Report circulation:

 * Public
 
# License

```{r ccby license, child=ggrParams$licenseCCBY}
```
 
# Citation

If you wish to use any of the material from this report please cite as:

 * Anderson, B. (`r 1900 + as.POSIXlt(Sys.Date())$year`) Processing report: NZ GREEN Grid project 1 minute electricity  power data, `r ggrParams$pubLoc`. XX replace with UKDA DOI when available XX

\newpage

# Introduction

```{r generic sample, child=ggrParams$sampleGeneric}
```
 

## Purpose

This report is intended to produce summary data quality statistics for the original GREEN grid Grid Spy household power demand monitoring data. This data was used to create a derived 'safe' dataset using the code in the `r ggrParams$repo` repository.

The cleaned 'safe' data has _no_ identifying information such as names, addresses, email addresses, telephone numbers and is therefore safe to share across all partners and to archive/release for research purposes via the UK Data Service's [ReShare](http://reshare.ukdataservice.ac.uk/) service.

The data contains a unique household id which can be used to link it to the anonymised NZ GREEN Grid household level data which will also be archived/release for research purposes via the UK Data Service's [ReShare](http://reshare.ukdataservice.ac.uk/) service.

## Requirements:

This report uses data quality statistics produced when processing the original grid spy 1 minute data downloads.

## History

Generally tracked via our github [repo](https://github.com/dataknut/nzGREENGridDataR):

 * [report history](https://github.com/dataknut/nzGREENGridDataR/commits/master/dataProcessing/gridSpy/buildGridSpy1mReport.Rmd)
 * [general issues](https://github.com/dataknut/nzGREENGridDataR/issues)
 
## Support


```{r generic support, child=ggrParams$supportGeneric}
```
 

# Original Data: Quality checks

The original data files files are stored on `r ggrParams$otagoHCS` at:

 * `r gSpyParams$gSpyInPath`

Data collection is ongoing and this section reports on the availability of data files collected up to the time at which the most recent safe file was created.

```{r getCompleteFileList}
# get file list 
fListCompleteDT <- data.table::as.data.table(readr::read_csv(gSpyParams$fListAll))

# for use below
nFiles <- nrow(fListCompleteDT)
nFilesNotLoaded <- nrow(fListCompleteDT[dateColName %like% "ignore"])
```

Overall we have `r tidyNum(nFiles)` files from `r tidyNum(uniqueN(fListCompleteDT$hhID))` households. However a large number of files contain no data as the monitoring devices have either been removed (households have moved or withdrawn from the study) or data transfer has failed.

Of the `r tidyNum(nFiles)`,  `r tidyNum(nFilesNotLoaded)` (`r round(100*(nFilesNotLoaded/nFiles),2)`%) were _not_ loaded/checked as their file sizes indicated that they contained no data.

## Date format checks

We now need to check how many of the loaded files have an ambiguous or default date - these could introduce errors.

```{r initialDateFormatTable}

# short cut if interim file list already saved ----
#ifile <- paste0(outPath, fListInterim)
#print(paste0("Loading 1 minute data files interim metadata to ", ifile))
#fListCompleteDT <- fread(ifile)
  
  
t <- fListCompleteDT[, .(nFiles = .N, 
                         minDate = min(dateExample), # may not make much sense
                         maxDate = max(dateExample)), 
                     keyby = .(dateColName, dateFormat)]

knitr::kable(t, 
             caption = "Number of files and min/max date (as char) with given date column names by inferred date format")
```

Results to note:

 * Following detailed checks there are now `r nrow(fListCompleteDT[dateFormat == "ambiguous"])` files with ambiguous dates;
 * The non-loaded files only have 2 distinct file sizes, confirming that they are unlikely to contain useful data. 
 
The following table lists up to 10 of the 'date NZ' files which are set by default - do they look OK to assume the default dateFormat? Compare the file names with the dateExample...

```{r listDefaultDateFilesNZT, echo=TRUE}
# list default files with NZ time
aList <- fListCompleteDT[dateColName == "date NZ" & dateFormat %like% "default", 
                         .(file, fSize, dateColName, dateExample, dateFormat)]

cap <- paste0("First 10 (max) of ", nrow(aList), 
              " files with dateColName = 'date NZ' and default dateFormat")

knitr::kable(caption = cap, head(aList))
```

The following table lists up to 10 of the 'date UTC' files which are set by default - do they look OK to assume the default dateFormat? Compare the file names with the dateExample...

```{r listDefaultDateFilesUTC, echo=TRUE}
# list default files with UTC time
aList <- fListCompleteDT[dateColName == "date UTC" & dateFormat %like% "default", 
                         .(file, fSize, dateColName, dateExample, dateFormat)]

cap <- paste0("First 10 (max) of ", nrow(aList), 
              " files with dateColName = 'date UTC' and default dateFormat")

knitr::kable(caption = cap, head(aList, 10))
```

Check final date formats:

```{r finalDateFormatTable}
# See what the date formats look like now
t <- fListCompleteDT[, .(nFiles = .N, 
                         minDate = min(dateExample), # may not make much sense
                         maxDate = max(dateExample)), 
                     keyby = .(dateColName, dateFormat)]

knitr::kable(t,
             caption = "Number of files & min/max dates (as char) with given date column names by final imputed date format")
```

## Data file quality checks

The following chart shows the distribution of these files over time using their sizes. Note that white indicates the presence of small files which may not contain observations.

```{r allFileSizesPlot}
myCaption <- paste0("Data source: ", gSpyParams$gSpyInPath,
                    "\nUsing data received up to ", max(fListCompleteDT$fMDate))

plotDT <- fListCompleteDT[, .(nFiles = .N,
                              meanfSize = mean(fSize)), 
                          keyby = .(hhID, date = as.Date(fMDate))]

ggplot2::ggplot(plotDT, aes( x = date, y = hhID, 
                             fill = log(meanfSize))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "black") + 
  scale_x_date(date_labels = "%Y %b", date_breaks = "1 month") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, hjust = 0.5)) + 
  labs(title = "Mean file size of all grid spy data files received per day",
       caption = paste0(myCaption, 
                        "\nLog file size used as some files are full year data")
    
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyAllFileListSizeTilePlot.png"))
```


The following chart shows the same chart but only for files which we think contain data.

```{r loadedFileSizesPlot}
plotDT <- fListCompleteDT[!is.na(dateFormat), .(nFiles = .N,
                              meanfSize = mean(fSize)), 
                          keyby = .(hhID, date = as.Date(fMDate))]

ggplot2::ggplot(plotDT, aes( x = date, y = hhID, fill = log(meanfSize))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "black") + 
  scale_x_date(date_labels = "%Y %b", date_breaks = "1 month") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  labs(title = "Mean file size of loaded grid spy data files received per day",
       caption = paste0(myCaption, 
                        "\nLog file size used as some files are full year data",
                        "\nFiles loaded if fsize > ", gSpyParams$gSpyFileThreshold, " bytes")
    
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyLoadedFileListSizeTilePlot.png"))
```


# Safe Data: quality checks

## Loaded data files

In this section we analyse the data files that have a file size > `r gSpyParams$gSpyFileThreshold` bytes and which have been used to create the safe data. Things to note:

 * We assume that any files smaller than this value have no observations. This is based on:
     * Manual inspection of several small files
     * The identical (small) file sizes involved
     * _But_ we should probably test the first few lines to double check...
 * We have had to deal with quite a lot of duplication some of which has caused the different date formats.
 
The following table shows the number of files per household that are actually processed to make the safe version.

```{r filesToLoadTable, echo=TRUE}
# check files to load
t <- fListCompleteDT[dateColName %like% "date", .(nFiles = .N,
                       meanSize = mean(fSize),
                       minFileDate = min(fMDate),
                       maxFileDate = max(fMDate)), keyby = .(hhID)]

knitr::kable(caption = "Summary of household files to load", t)
```


## Circuit label checks

The following table shows the number of households with different circuit labels by household. In theory each household should only have one set of unique circuit labels.

```{r circuitLabelCheck}
# get file list 
fLoadedListDT <- data.table::as.data.table(readr::read_csv(gSpyParams$fLoadedStats))

t <- fLoadedListDT[, .(nFiles = .N), keyby = .(hhID,circuitLabels)]

knitr::kable(caption = "Circuit labels list by number of files per household", t)

```

If this is not the case then this implies that:

 * some of the circuit labels for these households may have been changed during the data collection process;
 * some of the circuit labels may have character conversion errors which have changed the labels during the data collection process;
 * at least one file from one household has been saved to a folder containing data from a different household (unfortunately the raw data files do _not_ contain household IDs in the data or the file names which would enable checking/preventative filtering). This will be visible in the table if two households appear to share _exactly_ the same list of circuit labels.

Some or all of these may be true at any given time!

Errors are easy to spot in the following plot where a hhID spans 2 or more circuit labels.

```{r plotCircuitLabelIssuesAsTile, fig.height=8}
dt <- fLoadedListDT[!is.na(circuitLabels), 
                     .(nFiles = .N,
                       minObsDate = min(minDateTime), # helps locate issues in data
                       maxObsDate = max(maxDateTime),
                       minFileDate = min(mDateTime), # helps locate issues in files
                       maxFileDate = max(mDateTime),
                       nObs = sum(nObs)),
                     keyby = .(circuitLabels, hhID)] # ignore NA - it is files not loaded due to size thresholds

ggplot2::ggplot(dt, aes(y = hhID, x = circuitLabels, 
                       fill = nObs)) +
  geom_tile() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + 
  labs(title = "Circuit label counts for all loaded grid spy data",
       caption = paste0(myCaption,
                        "\nOnly files of size > ", 
                        gSpyParams$gSpyFileThreshold, " bytes loaded")
       
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyLoadedFileCircuitLabelsPlot.png"))
```

The following table provides more detail to aid error checking. Check for:

 * 2+ adjacent rows which have exactly the same circuit labels but different hh_ids. This implies some data from one household has been saved in the wrong folder;
 * 2+ adjacent rows which have different circuit labels but identical hh_ids. This could imply the same thing but is more likely to be errors/changes to the circuit labelling. 
 
If the above plot and this table flag a lot of errors then some re-naming of the circuit labels (column names) may be necessary. 

> NB: As before, the table is only legible in the html version of this report because latex does a very bad job of wrapping table cell text. A version is saved in `r paste0(gSpyParams$gSpyOutPath, "circuitLabelMetaDataCheckTable.csv")` for viewing in e.g. xl.

```{r circuitLabelMetaDataCheck}
knitr::kable(dt, 
             caption = "Circuit labels by household with addiitonal meta-data")

ofile <- paste0(gSpyParams$gSpyOutPath, "/checkStats/circuitLabelMetaDataCheckTable.csv")
write.csv(dt, ofile)
```

Things to note:

 * rf_25 has an aditional unexpected "Incomer 1 - Uncontrolled$2757" circuit in some files but it's value is always NA
 
## Observations

The following plots show the number of observations per day per household. In theory we should not see:

 * dates before 2014 or in to the future. These may indicate:
    * date conversion errors;
 * more than 1440 observations per day. These may indicate:
    * duplicate time stamps - i.e. they have the same time stamps but different power (W) values or different circuit labels;
    * observations from files that are in the 'wrong' rf_XX folder and so are included in the 'wrong' household as 'duplicate' time stamps.

If present both of the latter may have been implied by the table above and would have evaded the de-duplication filter which simply checks each complete row against all others within it's consolidated household dataset (a _within household absolute duplicate_ check).

```{r loadedFilesObsPlots}
# get file list 
hhStatDT <- data.table::as.data.table(readr::read_csv(gSpyParams$hhStatsByDate))
myCaption <- paste0(myCaption,
                        "\nOnly files of size > ", 
                        gSpyParams$gSpyFileThreshold, " bytes loaded")
  
# tile plot ----
ggplot2::ggplot(hhStatDT, aes( x = date, y = hhID, 
                               fill = nObs)) +
  geom_tile() +
  scale_fill_gradient(low = "red", high = "green") +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5,
                                   hjust = 0.5)) + 
  labs(title = "N observations per household per day for all loaded grid spy data",
       caption = myCaption
       
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyLoadedFileNobsTilePlot.png"))

# point plot ----
ggplot2::ggplot(hhStatDT, aes( x = date, 
                               y = nObs, 
                               colour = hhID)) +
  geom_point() +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, 
                                   hjust = 0.5)) + 
  labs(title = "N observations per household per day for all loaded grid spy data",
       caption = myCaption
       
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyLoadedFileNobsPointPlot.png"))
```

The following table shows the min/max observations per day and min/max dates for each household. As above, we should not see:

 * dates before 2014 or in to the future (indicates date conversion errors)
 * more than 1440 observations per day (indicates potentially duplicate observations)
 * non-integer counts of circuits as it suggests some column errors
 
 We should also not see NA in any row (indicates date conversion errors). 
 
 If we do see any of these then we still have data cleaning work to do!

```{r summaryTable}
# Stats table (so we can pick out the dateTime errors)
t <- hhStatDT[, .(minObs = min(nObs),
             maxObs = max(nObs), # should not be more than 1440, if so suggests duplicates
             meanN_Circuits =mean(nCircuits), #i.e. n circuits
             minDate = min(date),
             maxDate = max(date)),
         keyby = .(hhID)]

knitr::kable(caption = "Summary observation stats by hhID", t)
```


Finally we show the total number of households which we think are still sending data.

```{r liveDataHouseholds}
plotDT <- hhStatDT[, .(nHH = uniqueN(hhID)), keyby = .(date)]

# point plot ----
ggplot2::ggplot(plotDT, aes( x = date, y = nHH)) +
  geom_point() +
  scale_x_date(date_labels = "%Y %b", date_breaks = "6 months") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 0.5)) + 
  labs(title = "N live households per day for all loaded grid spy data",
       caption = myCaption
       
  )
ggplot2::ggsave(paste0(gSpyParams$gSpyOutPath, "/checkPlots/gridSpyLiveHouseholdsToDate.png"))

```

# Runtime


```{r check runtime, include=FALSE}
t <- proc.time() - startTime

elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * lubridate - date manipulation [@lubridate]
 * ggplot2 - for slick graphics [@ggplot2]
 * readr - for csv reading/writing [@readr]
 * dplyr - for select and contains [@dplyr]
 * progress - for progress bars [@progress]
 * knitr - to create this document & neat tables [@knitr]
 * kableExtra - for extra neat tables [@kableExtra]
 * nzGREENGrid - for local NZ GREEN Grid project utilities

Session info:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
